%%% A template for a simple PDF/A file like a stand-alone abstract of the thesis.

\documentclass[12pt]{report}

\usepackage[a4paper, hmargin=1in, vmargin=1in]{geometry}
\usepackage[a-2u]{pdfx}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{textcomp}

\begin{document}

%% Do not forget to edit abstract.xmpdata.

C\# is a strongly typed language utilizing type inference to save type annotations written by a programmer. 
However, the type inference is not as strong as other programming languages like Rust or Haskell.
This thesis aims to propose the C\# language improvement which would improve the current type inference and would be likely accepted by the C\# language design team. For this achievement, we analyzed Rust's type inference and observed necessary language requirements and type inference restrictions based on Hindley-Millner's formalization of type inference. 
These observations were used to propose a language improvement consisting of two parts. 
The first part is a C\# specification change describing the improvement by adjusting formal C\# specification. 
This part was presented to the language design team which resulted in a positive reaction where the team decided to keep moving forward with the proposal to make it available in the future C\# language version. 
The second part is an implementation of the improvement in the official C\# language compiler, Roslyn. 
The implementation is tested by using the original compiler tests and new tests testing the proposal functionality.  

\end{document}
