\chapter*{Conclusion}
\addcontentsline{toc}{chapter}{Conclusion}

The section \ref{sect:summary} of the thesis's introduction sets four goals, which are gradually achieved by the thesis. 
It starts by exploring C\# type inference by using the language specification and implementation in the Roslyn project. 
Then, the thesis compares C\# type inference with Rust type inference, which offers more advanced type inference than C\#. 
The type inference difference is explained using the theory background of Hindley-Milner type inference, which describes the limits of type inference in strongly typed languages.
\par
Based on these observations, the thesis selects a subset of type inference improvements suggested in C\# language discussions on GitHub and explores the motivation behind it to make it likely to add to the C\# language. 
The motivation is described in the created proposal, which is the fourth goal and which describes the improvement in terms of changing the specification.
\par
Together with the proposal, the implementation was made in the Roslyn project fork, with unit tests checking the functionality. 
This is the third and last goal of the thesis. 
The implementation ensured that basic compiler tests passed to reveal possible regression.
\par
In addition to the mentioned goals, the thesis succeeded in presenting the proposal and implementation to the LDM responsible for approving language changes into C\#. 
The committee agreed to continue with the change for further discussion and is generally inclined to ship it with future language releases.
\par
Based on the goals achieved and additional success in the LDM meeting, which was not a part of the scope of the thesis, we claim that the thesis accomplished all the promised goals.
